---
title: "Mini Data-Analysis Deliverable 1"
author: "April Hwang"
output: 
    github_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: cosmo
    highlight: tango
  
---

# Introduction

This is the Mini Data-Analysis Deliverable 1 for STATS 545A by April Hwang. It demonstrates my knowledge in using dplyr and ggplot2. 


### Load libraries, control message flow

```{r}
suppressPackageStartupMessages(library(tidyverse)) #hide package start up message

library(tidyverse) #load tidyverse
library(datateachr) #load datateachr

knitr::opts_chunk$set(message = FALSE) #hide messages
knitr::opts_chunk$set(warning = FALSE) #hide warnings

```

<br>

# Task 1: Choosing the dataset 

### 1.1 **(1 point)** Choose 4 datasets
<br>

1. `apt_buildings`
2. `parking_meters`
3. `steam_games`
4. `vancouver_trees`

<br>

### 1.2 **(6 points)** Find 3 attributes for each dataset 

<br>

#### **Attribute 1 - class type**
```{r}
class(apt_buildings)
class(parking_meters)
class(steam_games)
class(vancouver_trees)
```
<br>

#### **Attribute 2 - rows and columns**
```{r}
glimpse(apt_buildings)
glimpse(parking_meters)
glimpse(steam_games)
glimpse(vancouver_trees)
```

<br>

#### **Attribute 3 - count unique dataset in column**
```{r}
#count unique dataset in apt_buildings
apt_summary_tbl <- summarise(apt_buildings, across(c(id:no_barrier_free_accessible_units), n_distinct)) 

apt_long_summary_tbl <- apt_summary_tbl %>% pivot_longer(cols = everything(), names_to = 'variable_name', values_to = 'n_distinct')
print(apt_long_summary_tbl, n = 37)

#count unique dataset in parking_meters
parking_summary_tbl <- summarise(parking_meters, across(c(meter_head:meter_id), n_distinct))

parking_long_summary_tbl <- parking_summary_tbl %>% pivot_longer(cols = everything(), names_to = 'variable_name', values_to = 'n_distinct')
print(parking_long_summary_tbl, n = 22)


#count unique dataset in steam_games
games_summary_tbl <- summarise(steam_games, across(c(id:discount_price), n_distinct))

games_long_summary_tbl <- games_summary_tbl %>% pivot_longer(cols = everything(), names_to = 'variable_name', values_to = 'n_distinct')
print(games_long_summary_tbl, n = 21)


#count unique dataset in vancouver_trees
trees_summary_tbl <- summarise(vancouver_trees, across(c(tree_id:latitude), n_distinct))

trees_long_summary_tbl <- games_summary_tbl %>% pivot_longer(cols = everything(), names_to = 'variable_name', values_to = 'n_distinct')
print(trees_long_summary_tbl, n = 20)
```

<br>

### 1.3 **(1 point)** Choose 1 dataset and explain your choice

I have chosen `apt_buildings` as the columns and rows are labelled intuitively. I have eliminated the other datasets such as `parking_meters` which contain many code columns such as _r_mf_9a_6p_ and `vancouver_trees` which contain ambiguous columns such as _assigned_. More importantly, I am most interested in the information in this dataset and feel that I can think of four research questions as indicated in the learning objectives.

<br>

### 1.4 **(2 points)** Think of 1 research question for your dataset

One research question that I would like to explore with this dataset is whether attitude towards pets have changed over the years by looking into the relationship between _year_built_ and _pets_allowed_. I would look into a correlation plot to see if there has been an increase or decrease in the percentage of pet friendly buildings that were built over the years. This data can help to formulate a supporting argument for an assertion of whether there appears to be a change in attitude towards pet ownership.

<br>

# Task 2: Exploring your dataset 

### 2.1 & 2.2 **(12 points)** **(4 points)** Complete 4 exercises, and provide an explanation and comments for each exercise

<br>  

#### **Exercise 1 - plot the distribution of a numeric variable**

This this exercise, I will plot the distribution of the _year_built_ variable.

I chose the _year_built_ variable because the this is a variable which can be used to answer a number of different interesting research questions. For example, you can look at the correlation data between _year_built_ and another variable (e.g. _pets_allowed, no_of_accessible_parking_spaces, non-smoking_building_, etc.) to see if there is an emerging pattern which may indicate a sociological shift such as more positive attitude towards pets or negative attitude towards pet ownership, as previously discussed, as well as other factors such as consideration for accessibility needs and smoking.


```{r}
apt_buildings %>%
ggplot(aes(x = year_built)) + 
    geom_density()
```

<br>

#### **Exercise 2 - investigate number of missing values per variable and plot this** 

I chose this exercise because investigating how many missing values there are per variable is something that I have frequently  had to do in the past during data analysis. I used to use the excel "find" function, however, this is a more efficient way of accomplishing the same task. Additionally, it can be interesting to look into the implications of missing data, in that the amount of data available itself may be pertinent to a research question. 


```{r}
apply(X = is.na(apt_buildings), MARGIN = 2, FUN = sum)
```

RE: plotting - as per Lucy's message on Slack "I'm going to say that it's okay to answer this question by (1) investigating by summary and (2) to say what the plot would look like, i.e. you don't have to make the plot if you don't want to.":

The summary table shows that there are some columns without any missing data (e.g. _id, bike_parking, property_type, rsn_) whereas others have significant number of missing data (e.g. _amenities, prop_management_company_name_). The fact that some columns have more missing information than others can lend itself to a research question, in that it may imply that the data was difficult to attain, the information is generally deemed as of lower importance, or some form of ambiguity in the information being required to collect.

If the above summary table was to be plotted, it would have the columns in the x-axis and the number of missing data on the y-axis. The x-axis would be organized in an ascending manner (i.e. columns which have no missing information would be situated on the far left side and the column with the most number of missing information would be situated on the far right side). Various types of graphs, such scatter plot, line or even bar graph, would generally be suitable for this data.

<br>

#### **Exercise 3 - explore the relationship between 2 variables in a plot**

For this exercise, I will explore the relationship between _no_of_accessible_parking_spaces_ and _year_built_.

I chose this exercise because another potential research question that I can explore is looking into the relationship between time and number of accessible parking spaces. This may be pertinent information as part of an inquiry regarding availability of accessibility facilities for people with physical disabilities over time.


```{r}
apt_buildings %>%
ggplot(aes(x = year_built, y = no_of_accessible_parking_spaces)) +
    geom_point()
```

<br>

#### **Exercise 4 - make a new tibble with a subset of your data, with variables and observations that you are interested in exploring**

For this exercise, I will be making a new tibble with year_built and pets_allowed only, arrange by year.

I chose this exercise because, as mentioned above, one research question that I am interested in exploring is the relationship between the year that a building was built and allowance for pet to examine if there is a correlation between time and pet allowance. This new tibble will provide a quick glimpse of the relevant data and can be further manipulated in the future to make a graph.


```{r}
apt_buildings %>%
  select(year_built, pets_allowed) %>%
  arrange(year_built)
```

<br>

# Task 3: Choose research questions 

### **(4 points)** Write 4 research questions that you would like to explore in Milestone 2

<br>

**Research question 1** : What, if any, is the relationship between time and allowance of pet ownership in this dataset? Similarly, is there any relationship between time and availability of accessible parking?

**Research question 2** : Is there a relationship between the number of units and the number of storeys? 

**Research question 3** : What are the variables for which information is most readily available (i.e. no missing data) and least readily available (i.e. large amount of missing data)?

**Research question 4** : What are the top three most popular building amenities? 
