---
title: "Mini Data-Analysis Deliverable 1"
author: "April Hwang"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
    highlight: tango
---

# Introduction

This is the Mini Data-Analysis Deliverable 1 for STATS 545A by April Hwang. It demonstrates my knowledge in using dplyr and ggplot2. 


### Load libraries

```{r}
library(datateachr)
library(tidyverse)
```

<br>

# Task 1: Choosing the dataset 

### 1.1 **(1 point)** Choose 4 datasets
<br>

1. `apt_buildings`
2. `parking_meters`
3. `steam_games`
4. `vancouver_trees`

<br>

### 1.2 **(6 points)** Find 3 attributes for each dataset 

<br>

#### **Attribute 1 - class type**
```{r}
class(apt_buildings)
class(parking_meters)
class(steam_games)
class(vancouver_trees)
```
<br>

#### **Attribute 2 - rows and columns**
```{r}
glimpse(apt_buildings)
glimpse(parking_meters)
glimpse(steam_games)
glimpse(vancouver_trees)
```

<br>

#### **Attribute 3 - count unique dataset in column**
```{r}
summarise(apt_buildings, across(where(is.character), n_distinct))
summarise(parking_meters, across(where(is.character), n_distinct))
summarise(steam_games, across(where(is.character), n_distinct))
summarise(vancouver_trees, across(where(is.character), n_distinct))
```

<br>

### 1.3 **(1 point)** Choose 1 dataset and explain your choice

I have chosen `apt_buildings` as the columns and rows are labelled intuitively. I have eliminated the other datasets such as `parking_meters` which contain code columns such as _r_mf_9a_6p_ and `vancouver_trees` which contain ambiguous columns such as _assigned_. Also, I feel most interested in the information in this dataset and feel that I can think of four research questions as indicated in the learning objectives.

<br>

### 1.4 **(2 points)** Think of 1 research question for your dataset

One research question that I would like to explore with this dataset is whether attitude towards pets have changed over the years by looking into the relationship between _year_built_ and _pets_allowed_. I would look into a correlation plot to see if there has been an increase or decrease in the percentage of pet friendly buildings that were built over the years. This data can help to formulate a supporting argument for an assertion of whether there appears to be a change in attitude towards pet ownership.

<br>

# Task 2: Exploring your dataset 

### 2.1 & 2.2 **(12 points)** **(4 points)** Complete 4 exercises, and provide an explanation and comments for each exercise

<br>  

#### **Exercise 1 - plot the distribution of a numeric variable**

This this exercise, I will plot the distribution of the _year_built_ variable.

I chose the _year_built_ variable because the this is a variable which can be used to answer a number of different interesting research questions. For example, you can look at the correlation data between year_built and another variable (e.g. pets_allowed, no_of_accessible_parking_spaces, non-smoking_building, etc.) to see if there is an emerging pattern which may indicate a sociological shift such as more positive attitude towards pets or negative attitude towards pet ownership, as previously discussed, as well as other factors such as consideration for accessibility needs and smoking.


```{r}
ggplot(apt_buildings, aes(x = year_built)) + 
    geom_density()
```

<br>

#### **Exercise 2 - investigate number of missing values per variable and plot this** 

I chose this exercise because investigating how many missing values there are per variable is something that I have frequently  had to do in the past during data analysis. I used to use the excel "find" function, however, this is a more efficient way of accomplishing the same task.


```{r}
apt_buildings %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))

#find a way to plot this

```

<br>

#### **Exercise 3 - explore the relationship between 2 variables in a plot**

For this exercise, I will explore the relationship between _no_of_accessible_parking_spaces_ and _year_built_.

I chose this exercise because another potential research question that I can explore is looking into the relationship between time and number of accessible parking spaces. This may be pertinent information as part of an inquiry regarding availability of accessibility facilities for people with physical disabilities.


```{r}
apt_buildings %>%
ggplot( aes(x = year_built, y = no_of_accessible_parking_spaces)) +
    geom_point()
```

<br>

#### **Exercise 4 - make a new tibble with a subset of your data, with variables and observations that you are interested in exploring**

For this exercise, I will be making a new tibble with year_built and pets_allowed only, arrange by year.

I chose this exercise because, as mentioned above, one research question that I am interested in exploring the relationship between the year that a building was built and allowance for pet to examine if there is a correlation between time and pet allowance. This new tibble will provide a quick glimpse of the relevant data and can be further manipulated in the future to make a graph.


```{r}
apt_buildings %>%
  select(year_built, pets_allowed) %>%
  arrange(year_built)
```

<br>

# Task 3: Choose research questions 

### **(4 points)** Write 4 research questions that you would like to explore in Milestone 2

<br>

**Research question 1** : What, if any, is the relationship between time and attitude towards pet ownership? Use the data to support the assertion.

**Research question 2** : Is there a relationship between the number of units and the number of storeys?

**Research question 3** : What, if any, is the relationship between time and availability of accessible parking spaces? Use data to support the assertion.

**Research question 4** : What are the top three most popular building amenities? 
